---
title: "Supplementary code – Legendre et al. (2020)"
author: "Lucas Legendre"
date: "3/16/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
knitr::opts_knit$set(root.dir = '/Users/lucasl/Desktop/DataLegendreetal2020/')

# WARNING: edit the working directory above to your preferred folder
```

Compiled under R version 3.6.2 (2019-12-12)

This document details all analyses performed in R for the study:
Legendre, L. J., Rubilar-Rogers, D., Musser, G. M., Davis, S. N., Otero, R., Vargas, A. O., and Clarke, J. A. 2020. A giant soft-shelled egg from the Late Cretaceous of Antarctica. <i>Nature</i>.

For more information regarding the study, datasets, and analyses, please refer to the Supplementary Information of this paper. If you have any additional questions, feel free to email me at <lucasjlegendre@gmail.com>.


### Loading packages
```{r}
library(AICcmodavg)
library(ape)
library(caper)
library(dplyr)
library(evobiR)
library(ggplot2)
library(MPSEM)
library(nlme)
library(nortest)
library(phylopath)
library(phytools)
library(RColorBrewer)
```



## 1 – Lepidosaur dataset

This part requires the use of dataset 1 ("Dataset1-lepidosaurs.txt"; see also Supplementary Table 1 in Legendre et al., 2020), and lepidosaur phylogenetic tree ("Lepidosaurtree.trees.nex").

- Loading the tree:
```{r}
tree<-read.nexus("Lepidosaurtree.trees.nex")
```


### Phylogenetic generalized least squares (PGLS)

#### Egg volume ~ SVL (snout-vent length)

- Loading the dataset and removing species with missing data
```{r}
data2<-read.table("Dataset1-lepidosaurs.txt", header=T); data2V<-subset(data2, !is.na(V)&!is.na(SVL))
pruned.tree<-drop.tip(tree, setdiff(tree$tip.label, data2V$Species))
data<-read.table("Dataset1-lepidosaurs.txt", header=T, row.names = "Species"); dataV<-subset(data, !is.na(V)&!is.na(SVL))
```

- Best fit for alpha parameter in OU (Ornstein-Uhlenbeck) model
```{r}
alpha <- seq(0, 1, 0.1)
fit <- list()
form <- log(V)~log(SVL)
for (i in seq_along(alpha)) {
  cor <- corMartins(alpha[i], phy = pruned.tree, fixed = T)
  fit[[i]] <- gls(form, correlation = cor, data = data, na.action=na.exclude, method = "ML")
}
sapply(fit, logLik)
```

- Building PGLS models and running model selection using AICc
```{r}
BM<-gls(log(V)~log(SVL), data=dataV, correlation=corBrownian(phy=pruned.tree), method="ML")
OU<-gls(log(V)~log(SVL), data=dataV, correlation=corMartins(1, phy=pruned.tree, fixed=T), method="ML")
Lambda<-gls(log(V)~log(SVL), data=dataV, correlation=corPagel(1, phy=pruned.tree), method="ML")
EB<-gls(log(V)~log(SVL), data=dataV, correlation=corBlomberg(0.1, phy=pruned.tree), method="ML")
OLS<-gls(log(V)~log(SVL), data=dataV, method="ML")

Cand.models = list()
Cand.models[[1]] = BM
Cand.models[[2]] = OU
Cand.models[[3]] = Lambda
Cand.models[[4]] = EB
Cand.models[[5]] = OLS

Modnames = paste(c("BM", "OU", "Lambda", "EB", "OLS"), sep = " ")
aictab(cand.set = Cand.models, modnames = Modnames, sort = T)
summary(Lambda)
```

- Testing for normality
```{r}
plot(Lambda, resid(., type="n")~fitted(.), main="Normalized Residuals v Fitted Values",abline=c(0,0))
res<-resid(Lambda, type="n")
par(mar=c(5.1,4.1,4.1,2.1))
qqnorm(res)
qqline(res)
lillie.test(residuals(Lambda))
lillie.test(chol(solve(vcv(pruned.tree)))%*%residuals(Lambda))
```

- Removing outliers and building a new model without them
```{r}
which(res>4); which(res<(-2.5))
newdata<-dataV[-c(89,168,222),]
newdata2<-data2V[-c(89,168,222),]
newtree<-drop.tip(pruned.tree, setdiff(pruned.tree$tip.label, newdata2$Species))

Newlambda<-gls(log(V)~log(SVL), data=newdata, correlation=corPagel(1, phy=newtree), method="ML")
summary(Newlambda)

plot(Newlambda, resid(., type="n")~fitted(.), main="Normalized Residuals v Fitted Values",abline=c(0,0))
res<-resid(Newlambda, type="n")
par(mar=c(5.1,4.1,4.1,2.1))
qqnorm(res)
qqline(res)
lillie.test(residuals(Newlambda))
lillie.test(chol(solve(vcv(newtree)))%*%residuals(Newlambda))
```

- Estimating the pseudo R-squared using caper
```{r}
#newlambda:
datacomp<-comparative.data(phy=newtree, data=newdata2, names.col="Species")
pgls<-pgls(log(V)~log(SVL), data=datacomp, lambda="ML")
summary(pgls)

#lambda:
datacompl<-comparative.data(phy=pruned.tree, data=data2V, names.col="Species")
pgls<-pgls(log(V)~log(SVL), data=datacompl, lambda="ML")
summary(pgls)
```

<i>Not much difference between the two models, assumptions of normality are not met in either of them – we use <b>lambda</b> instead of <b>newlambda</b> for the plot.</i>

- Plot of the regression
```{r}
ggplot(dataV, aes(log(SVL), log(V), color=Suborder)) +
geom_point(size=4) +
  xlab("ln snout-vent length (mm)") +
  ylab("ln egg volume (mm3)") +
  geom_abline(intercept=-0.7990941, slope=1.6708400, colour="lightblue", size=1.3) +
  theme(panel.background = element_rect(fill="black")) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
        scale_colour_brewer("Suborder", palette="Accent")
```

- Estimating the SVL of <i>Antarcticoolithus</i> using the allometric equation of PGLS, including it in the dataset, then in the plot
```{r}
exp((-0.7990941/1.6708400)+(1/1.6708400)*log(5471405.789))

data[1,7]<-6684.303
dataV2<-subset(data, !is.na(V))

ggplot(dataV2, aes(log(SVL), log(V), color=Suborder)) +
  geom_point(size=4) +
  xlab("ln snout-vent length (mm)") +
  ylab("ln egg volume (mm3)") +
  geom_abline(intercept=-0.7990941, slope=1.6708400, colour="lightblue") +
  theme(panel.background = element_rect(fill="black")) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_colour_brewer("Suborder", palette="Accent")
```

<i>Eggshell thickness ~ body mass is not detailed here, since the code is virtually identical – only variable names and regression coefficients for the plot and estimation are different.</i>

- Estimating the body mass of <i>Antarcticoolithus</i> using the allometric equation of PGLS, including it in the dataset, then in the plot
```{r}
LambdaBM<-gls(log(V)~log(BM), data=dataV, correlation=corPagel(1, phy=pruned.tree), method="ML")
summary(LambdaBM)

exp(-(5.050557/0.619536)+(1/0.619536)*log(5471405.789))

data[1,8]<-21657214
dataV2<-subset(data, !is.na(V))

ggplot(dataV2, aes(log(BM), log(V), color=Suborder)) +
  geom_point(size=4) +
  xlab("ln body mass (g)") +
  ylab("ln egg volume (mm3)") +
  geom_abline(intercept=5.050556, slope=0.619536, colour="lightblue") +
  theme(panel.background = element_rect(fill="black")) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_colour_brewer("Suborder", palette="Set1")
```


#### Eggshell thickness ~ Egg volume

- Removing species with missing data
```{r}
dataVT2<-subset(data2, !is.na(Thickness)&!is.na(V)); dataVT2<-dataVT2[-1,]
pruned.treeVT<-drop.tip(tree, setdiff(tree$tip.label, dataVT2$Species))
dataVT<-subset(data, !is.na(Thickness)&!is.na(V)); dataVT<-dataVT[-1,]
```

- Building PGLS models and running model selection using AICc
```{r}
BMVT<-gls(log(Thickness)~log(V), data=dataVT, correlation=corBrownian(phy=pruned.treeVT), method="ML")
OUVT<-gls(log(Thickness)~log(V), data=dataVT, correlation=corMartins(1, phy=pruned.treeVT), method="ML")
LambdaVT<-gls(log(Thickness)~log(V), data=dataVT, correlation=corPagel(1, phy=pruned.treeVT), method="ML")
EBVT<-gls(log(Thickness)~log(V), data=dataVT, correlation=corBlomberg(0.1, phy=pruned.treeVT, fixed=T), method="ML")
OLSVT<-gls(log(Thickness)~log(V), data=dataVT, method="ML")

Cand.models = list()
Cand.models[[1]] = BMVT
Cand.models[[2]] = OUVT
Cand.models[[3]] = LambdaVT
Cand.models[[4]] = EBVT
Cand.models[[5]] = OLSVT

Modnames = paste(c("BMVT", "OUVT", "LambdaVT", "EBVT", "OLSVT"), sep = " ")
aictab(cand.set = Cand.models, modnames = Modnames, sort = T)
summary(EBVT)
```

- Testing for normality
```{r}
lillie.test(residuals(EBVT))
lillie.test(chol(solve(vcv(pruned.treeVT)))%*%residuals(EBVT))

plot(EBVT, resid(., type="n")~fitted(.), main="Normalized Residuals v Fitted Values",abline=c(0,0))
res<-resid(EBVT, type="n")
par(mar=c(5.1,4.1,4.1,2.1))
qqnorm(res)
qqline(res)
```


- Removing outliers and building a new model without them
```{r}
which(res<(-2)); which(res>2.5)
newdataVT<-dataVT[-c(19,45),]
newdataVT2<-dataVT2[-c(19,45),]
newtreeVT<-drop.tip(pruned.treeVT, c("Zootoca_vivipara_viviparous","Chamaeleo_senegalensis"))

NewEBVT<-gls(log(Thickness)~log(V), data=newdataVT, correlation=corBlomberg(0.1, phy=newtreeVT, fixed=T), method="ML")
summary(NewEBVT)

plot(NewEBVT, resid(., type="n")~fitted(.), main="Normalized Residuals v Fitted Values",abline=c(0,0))
res<-resid(NewEBVT, type="n")
par(mar=c(5.1,4.1,4.1,2.1))
qqnorm(res)
qqline(res)
lillie.test(residuals(NewEBVT))
lillie.test(chol(solve(vcv(newtreeVT)))%*%residuals(NewEBVT))
```

- Plot of the regression
```{r}
NewEBVT$coefficients
ggplot(newdataVT, aes(log(V), log(Thickness), color=Suborder)) +
  geom_point(size=4) +
  xlab("ln egg volume (mm3)") +
  ylab("ln eggshell thickness (µm)") +
  geom_abline(intercept=2.1246786, slope=0.3209485, colour="lightblue") +
  theme(panel.background = element_rect(fill="black")) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_colour_brewer("Suborder", palette="Accent")
```


### Phylogenetic path analysis

- Building the models
```{r}
ppadata<-dataVT
ppadata[,c(4,6:8)]<-log(ppadata[,c(4,6:8)])
M<-define_model_set(null=c(),
                    one=c(Thickness~SVL),
                    two=c(V~BM),
                    three=c(Thickness~V),
                    four=c(Thickness~V+SVL),
                    five=c(Thickness~V, V~BM),
                    six=c(Thickness~SVL, V~BM),
                    seven=c(Thickness~SVL+BM),
                    eight=c(V~BM, Thickness~BM),
                    nine=c(Thickness~V+BM),
                    ten=c(Thickness~SVL+BM+V),
                    eleven=c(Thickness~V+BM, V~BM),
                    twelve=c(Thickness~SVL+BM, V~BM),
                    thirteen=c(Thickness~BM),
                    .common=c(BM~SVL, V~SVL))
```

- Plot all models and compile best model using CICc
```{r}
plot_model_set(M)
result<-phylo_path(M, data=ppadata, tree=pruned.treeVT, model='lambda'); result
s<-summary(result); s
plot(s)
averagemodel<-average(result)
plot(averagemodel)
```


### Predictions of SVL and body mass – using phylogenetic eigenvectors maps (PEM)

- Data and tree
```{r}
fossiltreePEM<-read.nexus("Lepidosaurtree.trees.nex")
fossildataPEM<-read.table("Dataset1-lepidosaurs.txt",header=TRUE,stringsAsFactor=FALSE,)
fossildataV<-subset(fossildataPEM, !is.na(V))
fossildataV[,c(3:5,7:9)]<-log(fossildataV[,c(3:5,7:9)])
```

#### For SVL

- Match the phylogeny and dataset
```{r}
treePEM<-drop.tip(fossiltreePEM, setdiff(fossiltreePEM$tip.label, fossildataV$Species))
treedrop <- drop.tip(treePEM,fossildataV[is.na(fossildataV[,"SVL"]),"Species"])
grloc <- getGraphLocations(treePEM,fossildataV[is.na(fossildataV[,"SVL"]),"Species"])
sporder <- match(attr(grloc$x,"vlabel")[grloc$x$vertex$species],fossildataV[,"Species"])
```

- Build the PEM
```{r}
PEMfs <- list()
PEMfs[["V"]] <- PEM.fitSimple(y=fossildataV[sporder,"SVL"],
                    x=fossildataV[sporder,"V"],w=grloc$x,d="distance",sp="species",
                    lower=0,upper=1)
PEMfs[["none"]] <- PEM.fitSimple(y=fossildataV[sporder,"SVL"],
                    x=NULL,w=grloc$x,d="distance",sp="species",lower=0,upper=1)
for(m in c("V","none")) print(PEMfs[[m]]$optim$par)
rm(m)
```

- Select the best model (with V as a co-predictor vs without co-predictors) based on AICc
```{r}
PEMAIC <- list()
PEMAIC[["V"]] <- lmforwardsequentialAICc(y=fossildataV[sporder,"SVL"],
                                         x=fossildataV[sporder,"V",drop=FALSE],object=PEMfs[["V"]])
PEMAIC[["none"]] <- lmforwardsequentialAICc(y=fossildataV[sporder,"SVL"],object=PEMfs[["none"]])
for(m in c("V","none"))
  cat(m,summary(PEMAIC[[m]])$adj,PEMAIC[[m]]$AICc,"\n")
rm(m)

summary(PEMAIC[["V"]])
```

<i>The model with V as a co-predictor has the best fit.</i>

- Predicting missing values for SVL
```{r}
tf<-list(V=identity, none=identity)
m <- "V" ; atr <- tf[[m]](fossildataV[is.na(fossildataV[,"SVL"]),m,drop=FALSE])
resultsPEM<-predict(object=PEMfs[[m]],targets=grloc,lmobject=PEMAIC[[m]],newdata=atr,interval="confidence")

# Predicted SVL, with upper and lower limits of the confidence interval for the prediction:
exp(resultsPEM$values); exp(resultsPEM$upper); exp(resultsPEM$lower)
```

<i>The code to predict missing values for body mass is identical to the one for SVL; simply replace 'SVL' with 'BM'.</i>

- Plotting the regression for Egg volume ~ SVL with the value of SVL for <i>Antarcticoolithus</i> estimated from PEM
```{r}
data[1,7]<-6665.867
dataV2<-subset(data, !is.na(V))

ggplot(dataV2, aes(log(SVL), log(V), color=Suborder)) +
  geom_point(size=4) +
  xlab("ln snout-vent length (mm)") +
  ylab("ln egg volume (mm3)") +
  geom_abline(intercept=-0.7990941, slope=1.6708400, colour="lightblue") +
  theme(panel.background = element_rect(fill="black")) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_colour_brewer("Suborder", palette="Accent")
```

- Plotting the regression for Egg volume ~ Body mass with the value of body mass for <i>Antarcticoolithus</i> estimated from PEM
```{r}
data[1,8]<-456096.7
dataV2<-subset(data, !is.na(V))

ggplot(dataV2, aes(log(BM), log(V), color=Suborder)) +
  geom_point(size=4) +
  xlab("ln body mass (g)") +
  ylab("ln egg volume (mm3)") +
  geom_abline(intercept=5.050556, slope=0.619536, colour="lightblue") +
  theme(panel.background = element_rect(fill="black")) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_colour_brewer("Suborder", palette="Set1")
```


### Phylogenetic ANOVA

- Subsetting the data and tree, and compiling relative eggshell thickness (RT)
with RT = (eggshell thickness / egg length)
```{r}
RT<-data$Thickness/data$L; data<-cbind(data, RT)
dataA<-subset(data, !is.na(RT)&!is.na(OV))
dataA$RT<-log(dataA$RT)
anovatree<-drop.tip(tree, setdiff(tree$tip.label, rownames(dataA)))
```

#### ANOVA with 2 groups (oviparous and viviparous)

- Subsetting the dataset and performing the ANOVA
```{r}
datANOVA<-as.data.frame(cbind(dataA$RT, dataA$OV), header=T)
row.names(datANOVA)<-row.names(dataA); colnames(datANOVA)<-c("RT", "OV")
datANOVA<-ReorderData(anovatree, datANOVA, taxa.names="row names")

phylANOVA(anovatree, datANOVA$OV, datANOVA$RT, p.adj="BH")
```

- Boxplot of both groups
```{r}
ggplot(dataA, aes(x=OV, y=RT, fill=OV)) +
  geom_boxplot() +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_fill_hue()

log(data$RT)[1] # Value for Antarcticoolithus, intermediate between the groups
```

- Extract percentiles for each boxplot
```{r}
p2<-ggplot(dataA, aes(x=OV, y=RT, fill=OV)) + geom_boxplot()
ggplot_build(p2)$data
```

#### ANOVA with 3 groups (oviparous and laid, oviparous and dissected from the oviduct, and viviparous)

- Subsetting the dataset and performing the ANOVA
```{r}
datANOVA2<-as.data.frame(cbind(dataA$RT, dataA$Strategy), header=T)
row.names(datANOVA2)<-row.names(dataA); colnames(datANOVA2)<-c("RT", "Strategy")
datANOVA2<-ReorderData(anovatree, datANOVA2, taxa.names="row names")

phylANOVA(anovatree, datANOVA2$Strategy, datANOVA2$RT, p.adj="BH")
```

- Boxplot of all groups
```{r}
ggplot(dataA, aes(x=Strategy, y=RT, fill=Strategy)) +
  geom_boxplot() +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_fill_hue()
```

- Extract percentiles for each boxplot
```{r}
p2<-ggplot(dataA, aes(x=Strategy, y=RT, fill=Strategy)) + geom_boxplot()
ggplot_build(p2)$data
```


## 2 – Amniote dataset

This part requires the use of dataset 2 ("Dataset2-amniotes.txt"; see also Supplementary Table 3 in Legendre et al., 2020), and amniote phylogenetic tree ("Amniotetree.nex").

- Loading and matching the data and tree
```{r}
dataS<-read.table("Dataset2-amniotes.txt", header=T)
treeS<-read.nexus("Amniotetree.nex")
treeS<-drop.tip(treeS, setdiff(treeS$tip.label, dataS$Taxon))
dataS<-read.table("Dataset2-amniotes.txt", header=T, row.names="Taxon")
Ws<-diag(vcv.phylo(treeS))
```


### PGLS – Eggshell thickness ~ Egg mass

#### For hard-shelled eggs

- Subset data and tree
```{r}
dataS<-read.table("Dataset2-amniotes.txt", header=T)
datahard<-dataS[-c(50,86:88,93:95,115,118,120:145,147,149),]
treehard<-drop.tip(treeS, setdiff(treeS$tip.label, datahard$Taxon))
dataS<-read.table("Dataset2-amniotes.txt", header=T, row.names="Taxon")
datahard<-dataS[-c(50,86:88,93:95,115,118,120:145,147,149),]
Wh<-diag(vcv.phylo(treehard))
```

- Best fit for alpha parameter in OU model
```{r}
alpha <- seq(0, 1, 0.1)
fit <- list()
form <- log(Eggshell_thickness)~log(Egg_mass)
for (i in seq_along(alpha)) {
  cor <- corMartins(alpha[i], phy = treehard, fixed = T)
  fit[[i]] <- gls(form, correlation = cor, data = datahard, na.action=na.exclude, weights=varFixed(~Wh), method = "ML")
}
sapply(fit, logLik)
```

- Best fit for g parameter in EB (Early Burst) model
```{r}
g <- seq(0.1, 1, 0.1)
fit <- list()
form <- log(Eggshell_thickness)~log(Egg_mass)
for (i in seq_along(g)) {
  cor <- corBlomberg(g[i], phy = treehard, fixed = T)
  fit[[i]] <- gls(form, correlation = cor, data = datahard, na.action=na.exclude, weights=varFixed(~Wh), method = "ML")
}
sapply(fit, logLik)
```

- Building PGLS models and running model selection using AICc
```{r}
BM<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datahard, correlation=corBrownian(phy=treehard), weights=varFixed(~Wh), method="ML")
OU<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datahard, correlation=corMartins(0.1, phy=treehard, fixed=T), weights=varFixed(~Wh), method="ML")
Lambda<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datahard, correlation=corPagel(1, phy=treehard), weights=varFixed(~Wh), method="ML")
EB<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datahard, correlation=corBlomberg(0.1, phy=treehard, fixed=T), weights=varFixed(~Wh), method="ML")
OLS<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datahard, method="ML")

Cand.models = list()
Cand.models[[1]] = BM
Cand.models[[2]] = OU
Cand.models[[3]] = Lambda
Cand.models[[4]] = EB
Cand.models[[5]] = OLS

Modnames = paste(c("BM", "OU", "Lambda", "EB", "OLS"), sep = " ")
aictab(cand.set = Cand.models, modnames = Modnames, sort = T)
summary(Lambda)
```

- Estimating the pseudo R-squared using caper
```{r}
dataS<-read.table("Dataset2-amniotes.txt", header=T)
datahard<-dataS[-c(50,86:88,93:95,115,118,120:145,147,149),]
datacomp<-comparative.data(phy=treehard, data=datahard, names.col="Taxon")
pgls<-pgls(log(Eggshell_thickness)~log(Egg_mass), data=datacomp, lambda="ML")
summary(pgls)
```

#### For soft-shelled eggs

- Subset data and tree
```{r}
datasoft<-dataS[c(50,86:88,93,94,115,118,120:145,147,149),]
treesoft<-drop.tip(treeS, setdiff(treeS$tip.label, datasoft$Taxon))
dataS<-read.table("Dataset2-amniotes.txt", header=T, row.names="Taxon")
datasoft<-dataS[c(50,86:88,93,94,115,118,120:145,147,149),]
Ws<-diag(vcv.phylo(treesoft))
```

- Best fit for alpha parameter in OU model
```{r}
alpha <- seq(0, 1, 0.1)
fit <- list()
form <- log(Eggshell_thickness)~log(Egg_mass)
for (i in seq_along(alpha)) {
  cor <- corMartins(alpha[i], phy = treesoft, fixed = T)
  fit[[i]] <- gls(form, correlation = cor, data = datasoft, na.action=na.exclude, weights=varFixed(~Ws), method = "ML")
}
sapply(fit, logLik)
```

- Best fit for g parameter in EB model
```{r}
g <- seq(0.1, 1, 0.1)
fit <- list()
form <- log(Eggshell_thickness)~log(Egg_mass)
for (i in seq_along(g)) {
  cor <- corBlomberg(g[i], phy = treesoft, fixed = T)
  fit[[i]] <- gls(form, correlation = cor, data = datasoft, na.action=na.exclude, weights=varFixed(~Ws), method = "ML")
}
sapply(fit, logLik)
```

- Building PGLS models and running model selection using AICc
```{r}
BM<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datasoft, correlation=corBrownian(phy=treesoft), weights=varFixed(~Ws), method="ML")
OU<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datasoft, correlation=corMartins(1, phy=treesoft), weights=varFixed(~Ws), method="ML")
Lambda<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datasoft, correlation=corPagel(0.2, phy=treesoft), weights=varFixed(~Ws), method="ML")
EB<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datasoft, correlation=corBlomberg(0.1, phy=treesoft, fixed=T), weights=varFixed(~Ws), method="ML")
OLS<-gls(log(Eggshell_thickness)~log(Egg_mass), data=datasoft, method="ML")

Cand.models = list()
Cand.models[[1]] = BM
Cand.models[[2]] = OU
Cand.models[[3]] = Lambda
Cand.models[[4]] = EB
Cand.models[[5]] = OLS

Modnames = paste(c("BM", "OU", "Lambda", "EB", "OLS"), sep = " ")
aictab(cand.set = Cand.models, modnames = Modnames, sort = T)
summary(Lambda)
```

- Estimating the pseudo R-squared using caper
```{r}
dataS<-read.table("Dataset2-amniotes.txt", header=T)
datasoft<-dataS[c(50,86:88,93,94,115,118,120:145,147,149),]
datacomp<-comparative.data(phy=multi2di(treesoft), data=datasoft, names.col="Taxon")
pgls<-pgls(log(Eggshell_thickness)~log(Egg_mass), data=datacomp, lambda="ML")
summary(pgls)
```

#### Plot with both regression lines (hard-shelled and soft-shelled)

```{r}
dataS<-read.table("Dataset2-amniotes.txt", header=T)
ggplot(dataS, aes(log(Egg_mass), log(Eggshell_thickness), colour=Clade)) +
  geom_point(size=4) +
  xlab("ln egg mass (g)") +
  ylab("ln calcareous layer thickness (µm)") +
  geom_abline(intercept=4.321066, slope=0.439690, colour="turquoise", size=1.3) +
  geom_abline(intercept=2.1224777, slope=0.3930565, colour="red", size=1.3) +
  theme(panel.background = element_rect(fill="black")) +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_colour_brewer("Group", palette="Set1")
```


### Ancestral state reconstruction 

- For ratio (Calcareous layer thickness / Egg mass), using 'fastAnc'
```{r}
dataS<-read.table("Dataset2-amniotes.txt", header=T); dataS<-dataS[-95,]
treeplot<-drop.tip(treeS, setdiff(treeS$tip.label, dataS$Taxon))
dataS<-read.table("Dataset2-amniotes.txt", header=T, row.names="Taxon"); dataS<-dataS[-95,]
dataplot=as.matrix(dataS[,3]); names(dataplot)<-rownames(dataS); colnames(dataplot)<-"CL_thickness"
dataplot<-log10(dataplot)

fit<-fastAnc(treeplot,dataplot,vars=T,CI=T) # Ancestral states for each node
obj<-contMap(treeplot,dataplot,plot=F)

plot(setMap(obj, colors=rev(brewer.pal(10, "Spectral"))), fsize=0.4,lwd=4)
plot(setMap(obj, colors=rev(brewer.pal(10, "Spectral"))), type="fan", fsize=0.4,lwd=4)
```

- Same as above, but with an ultrametric tree, as shown in Fig. 3 of Legendre et al. (2020)
```{r}
treeplot<-force.ultrametric(treeplot)
fit<-fastAnc(treeplot,dataplot,vars=T,CI=T) # Ancestral states for each node
obj<-contMap(treeplot,dataplot,plot=F)

plot(setMap(obj, colors=rev(brewer.pal(10, "Spectral"))), fsize=0.4,lwd=4)
```

- For presence/absence of the prismatic layer, using 'make.simmap'
```{r}
dataS<-read.table("Dataset2-amniotes.txt", header=T)
dataS<-dataS[-95,]
treeS<-read.nexus("Amniotetree.nex")
treeS<-drop.tip(treeS, setdiff(treeS$tip.label, dataS$Taxon))
prismatic<-dataS[,5]; names(prismatic)<-dataS$Taxon
cols<-setNames(c("royalblue","red3"),levels(prismatic))

mtree<-make.simmap(treeS, prismatic, model="ER")
plot(mtree,cols,type="fan",fsize=0.5,ftype="i")
add.simmap.legend(colors=cols,prompt=FALSE,x=0.9*par()$usr[1],
                  y=0.8*par()$usr[3],fsize=0.8)

# with 1000 simulations of stochastic character maps from the data:
mtrees<-make.simmap(treeS, prismatic, model="ER", nsim=1000)
par(mfrow=c(10,10))
null<-sapply(mtrees,plotSimmap,colors=cols,lwd=1,ftype="off")
pd<-summary(mtrees); pd
par(mfrow=c(1,1))
plot(pd,fsize=0.6,ftype="i",colors=cols,ylim=c(-2,Ntip(treeS)))
add.simmap.legend(colors=cols[2:1],prompt=FALSE,x=0,y=-2,vertical=FALSE)

# density map of the above
objprism<-densityMap(mtrees,states=levels(prismatic)[2:1],plot=FALSE)
n<-length(objprism$cols)
objprism$cols[1:n]<-colorRampPalette(c("royalblue","red3"))(n)
plot(objprism,size=c(0.6,1),fsize=0.6,lwd=4)
```


### Boxplot for ratio (Calcareous layer thickness / Egg mass)

```{r}
dataS<-read.table("Dataset2-amniotes.txt", header=T)
RTM<-dataS$Eggshell_thickness/dataS$Egg_mass
dataS<-cbind(dataS, RTM)
dataSboxplot<-dataS %>% filter(!Clade=="Test")

ggplot(dataSboxplot, aes(x=Clade, y=log(RTM), fill=Clade)) + 
  geom_boxplot() +
  theme(panel.background = element_rect(fill="white")) +
  theme(axis.text=element_text(size=10, angle = 90),
        axis.title=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=17)) +
  scale_fill_brewer(palette="Set1")
```

- Extract percentiles for each boxplot
```{r}
p<-ggplot(dataSboxplot, aes(x=Clade, y=log(RTM), fill=Clade)) + 
  geom_boxplot()
ggplot_build(p)$data
```
